---
output:
  html_document:
    highlight: null
    self_contained: no
    theme: null
---

```{r define-funcs, echo=FALSE, warning=FALSE, message=FALSE}
source("code-plot_git.R")
```

Yesterday I saw [this retweet](https://t.co/u3uHy6a9d4) from [\@timelyportfolio](https://twitter.com/timelyportfolio) that links to a [gist by \@gka](https://gist.github.com/gka/393f5ab2b95e927d305eb6e14767180e).

The gist gave a few short scripts that can show you when your git commits take place during the day. I thought that was cool, so I took it a step (or five) further by writing it all up in R code and made nice wrapper functions for it and added interactive visualizations and made a shiny app... I just had a bit of fun. This is the result.

Note that all commit times are reported as PST timezone.

**TL;DR: I'm dead before 10am --- My prime working time is 2am --- My supervisor Jenny Bryan has a very compatible schedule to mine --- Hadley somehow learned in the past 3 years to be so productive that he doesn't even work on weekends**

## **My work hours since first learning R**

This plot shows the times of all my git commits since Sept 2013, colour-coded by git repo (aka project).

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=12}
library(plotly)
plot_git_commits(logfile = "data/dean-projects.csv", date_begin = "2013-09-01",
                 plot_type = "ggplot")
```

September 2013 is when I first learned R in the beginning of my masters degree, and the project I was working on then ([statsTerrorismProject](https://github.com/daattali/statsTerrorismProject)) was my final project for STAT545.

After taking my introductory R course, I spent the next year on my other masters courses and didn't do any R-ing and coding, hence the giant gap. The next time I used git was when I took the second half of STAT545 in 2014, and that's where I learned how to write a package -- [rsalad](https://github.com/daattali/rsalad) was my first R package, developed as homework. Then I didn't do any open source work for a few months because I was working in a lab that wasn't very supportive of that... And in January 2015 I started working with Jenny as my supervisor, and as you can see, I was pretty much busy with just about everything in the world except for my actual masters project! :)

#### **A few things that are easy to notice:**

- I refuse to work before 10am
- It looks like I work until 2am-4am fairly consistently
- I did a LOT of work on my website ([dattali.github.io](http://daattali.github.io)) around March 2015 (which is when I got the idea to make a website and when I released it)
- There were a few days in July 2015 where I started coding at 7am! Nope, that's a lie, I was in Toronto that week so that's actually 10am as usual...
- It's cool to see when I work on which projects. You can very clearly see the two-week effort on [timevis](https://github.com/daattali/timevis) in July 2016 for example

## **My work hours in the past 6 months** (Feb-Aug 2016)

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=12}
plot_git_commits(logfile = "data/dean-projects.csv", date_begin = "2016-02-18",
                 date_end = "2016-08-18", plot_type = "ggplot")
```

#### **Observations**

It's interesting to see how my commits very closely relate to what's going on in my life :)

- During Feb-March I was writing my masters thesis on my package [ddpcr](https://github.com/daattali/ddpcr) and you can see how I had one last burst of work on the package at the end of Feb, and then I completely stopped coding for a few weeks while I focused on writing. I was such a good boy.
- June seems like a dead month. I was in Toronto, Berkeley and Stanford for conferences and didn't code at all. It's cool how clearly that shows up!
- It's nice to see that [shinyjs](https://github.com/daattali/shinyjs) has reached a fairly stable state and I don't have to spend much time on it anymore

## **Adding marginal density plots to see exactly what times are alive/dead**

Just for fun, I can use my [ggMarginal()](https://github.com/daattali/ggExtra) function to add marginal density plots, to make it more clear when most of my commits take place.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=12}
plot_git_commits(logfile = "data/dean-projects.csv", date_begin = "2016-02-18",
                 date_end = "2016-08-18", plot_type = "density")
```

## **How's my (ex) supervisor Jenny Bryan doing?**

After working with Jenny for a couple years, I learned that we have very compatible schedules. It was very normal for us to exchange emails past midnight when we had a 9am class the following morning, and I'd often see her making commits at 1-2am when I was just getting in the zone!

With this graph we can see that our schedules really are very similar, although she is shifted 2 hours from me: I work 10am-4am, she works 8am-2am. What an early bird.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=12}
plot_git_commits(logfile = "data/jenny-projects.csv", date_begin = "2010-01-01",
                 plot_type = "ggplot")
```

## **And the grand finale: the R master Hadley**

No introduction needed, the name speaks for itself. (Note these are all PST times)

Just one thing I'd like to note: we've all heard "this is going to be the year of 'ggvis'!", yet it looks like 2013 was the year-est of 'ggvis' so far :)

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=12}
plot_git_commits(logfile = "data/hadley-projects.csv", date_begin = "2010-01-01",
                 plot_type = "ggplot")
```

## **Working on weekends**

If you look at day of the week, you'll see that weekends and weekdays are essentially the same to me.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=12}
plot_git_commits(logfile = "data/dean-projects.csv", date_begin = "2010-01-01",
                 plot_type = "ggplot", x = "date", y = "weekday")
```

But what's amazing is that if you look at Hadley, he seems to be such a good time manager that he manages to almost never work on weekends in the past 3 years, despite producing output at a speed of a dozen adults. That discipline of not working on weekends is commendable.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=12}
plot_git_commits(logfile = "data/hadley-projects.csv", date_begin = "2010-01-01",
                 plot_type = "ggplot", x = "date", y = "weekday")
```